---
title: "Laing-Bergelson-Infancy"
author: "Catherine Laing & Elika Bergelson"
date: "July 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(lme4)
library(effsize)

source("Demographics.R")

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

See Table 1 for age information

```{r Table 1 - Mean age}
read_csv("data/17_month_ages.csv") %>%
  summarise(mean_age_CDI = mean(days_CDI, na.rm=T), 
            sd_age_CDI = sd(days_CDI, na.rm=T),
            mean_age_video = mean(days_video, na.rm=T),
            sd_age_video = sd(days_video, na.rm=T),
            mean_age_audio = mean(days_audio, na.rm=T),
            sd_age_audio = sd(days_audio, na.rm=T),
            mean_age_demo18 = mean(days_demo18, na.rm=T),
            sd_age_demo18 = sd(days_demo18, na.rm=T),
            mean_age_demo6 = mean(days_demo6, na.rm=T),
            sd_age_demo6 = sd(days_demo6, na.rm=T))
```


FT mothers worked an average of 38 hours/week (Range: 30-45, SD=4.4) and PT mothers worked an average of 18.5 hours/week (Range=6-25, SD=6). 
```{r Summary of work status groups}
workdata %>%  
  group_by(MOTwork18) %>%
  summarize(mean_MOTwork18 = mean(MOTworkhours18, na.rm=T),
           mdn_MOTwork18 = median(MOTworkhours18, na.rm=T),
           sd_MOTwork18 = sd(MOTworkhours18, na.rm=T),
           min_MOTwork18 = min(MOTworkhours18, na.rm=T),
           max_MOTwork18 = max(MOTworkhours18, na.rm=T))

workdata %>%  
  group_by(MOTwork6) %>%
  summarize(mean_MOTwork6 = mean(MOTworkhours6, na.rm=T),
           mdn_MOTwork6 = median(MOTworkhours6, na.rm=T),
           sd_MOTwork6 = sd(MOTworkhours6, na.rm=T),
           min_MOTwork6 = min(MOTworkhours6, na.rm=T),
           max_MOTwork6 = max(MOTworkhours6, na.rm=T))
```


```{r Caretype groups}
workdata <- workdata %>%
   mutate(caretype = ifelse(MOTwork18 == "Full-time" & MOTwork6 == "Full-time", "Care-only", "Other"),
          caretype = ifelse(MOTwork18 == "Full-time" & MOTwork6 == "Part-time", "Mixed-care-nonmaternal", caretype),
          caretype = ifelse(MOTwork18 == "Full-time" & MOTwork6 == "Home", "Mixed-care-maternal", caretype),
          caretype = ifelse(MOTwork18 == "Part-time" & MOTwork6 == "Part-time", "Mixed-care-equal", caretype),
          caretype = ifelse(MOTwork18 == "Part-time" & MOTwork6 == "Home", "Mixed-care-maternal2", caretype),
          caretype = ifelse(MOTwork18 == "Home" & MOTwork6 == "Home", "Home-only", caretype))

```

```{r Caretype group size}

workdata %>% group_by(caretype, sex) %>% tally()

workdata <- workdata %>% mutate(caretype = fct_recode(caretype, 
                                                      "Mixed-care" = "Mixed-care-nonmaternal",
                                                      "Mixed-care" = "Mixed-care-maternal",
                                                      "Mixed-care" = "Mixed-care-equal",
                                                      "Mixed-care" = "Mixed-care-maternal2",
                                                      "Home-only" = "Other"))

write_csv(workdata, "data/workdata.csv") # save for Bayes analysis

```

Five infants' (2 females) CDI data was not available, but audio and video data were recorded for all 44 infants. Shapiro-Wilkes tests revealed that all three vocabulary measures (CDI, types, and tokens) were not normally distributed. They were thus log-transformed (after adding one, since four infants produced zero words in the recorded data; reported vocabulary ranged from 1-372 words; M=60.44, SD=76.3). 

```{r Summary of data}

workdata %>% filter(noCDI==T) # CDIs unavailable for 5 infants

workdata %>%
  mutate(nowords = ifelse(CDI == 0, T, F)) %>%
  filter(nowords == TRUE) # 0 infants + 5 with no cdi at all

workdata %>%
  mutate(nowords = ifelse(CHItokens == 0, T, F)) %>%
  filter(nowords == TRUE) # 5 infants

workdata %>%
  summarize(meanProd = mean((CDI), na.rm=T),
           sdProd = sd((CDI), na.rm=T),
           maxProd = max(CDI, na.rm=T),
           minProd = min(CDI, na.rm=T))
```

Pearson product-moment correlations on the log-transformed data revealed significant correlations between our three measures (see Table 3), substantiating the parental report data.

```{r Correlation between variables - table 3}
cor.test(workdata$Log.CDI, workdata$Log.CDInouns, method = 'pearson') # significant, p<.001
cor.test(workdata$Log.CDI, workdata$Log.Types, method = 'pearson') # significant, p<.001
cor.test(workdata$Log.CDI, workdata$Log.Tokens, method = 'pearson') # significant, p<.001

cor.test(workdata$Log.CDInouns, workdata$Log.Types, method = 'pearson') # significant, p<.001
cor.test(workdata$Log.CDInouns, workdata$Log.Tokens, method = 'pearson') # significant, p<.001

cor.test(workdata$Log.Types, workdata$Log.Tokens, method = 'pearson') # significant, p<.001
```

```{r Testing CDI data}
shapiro.test(workdata$CDI) # not normal
shapiro.test(workdata$Log.CDI) # normal

workdata <- workdata %>%
  mutate(outlier.CDI = (Log.CDI > 
                      (mean(Log.CDI, na.rm=T) + 
                         3*(sd(Log.CDI, na.rm=T))) |
                      Log.CDI <
                      (mean(Log.CDI, na.rm=T) - 
                         3*(sd(Log.CDI, na.rm=T)))))

workdata %>% filter(outlier.CDI == T) # no outliers

workdata %>% summarize(M = mean(CDI, na.rm=T), 
                                              SD = sd(CDI, na.rm=T))

workdata %>% group_by(caretype) %>% summarize(M = mean(CDI, na.rm=T), 
                                              SD = sd(CDI, na.rm=T))

mod1cdi<- lm(Log.CDI~caretype, workdata)
mod2cdi<- lm(Log.CDI~caretype+sex+MOTedu, workdata)
mod3cdi<- lm(Log.CDI~caretype+sex+PVTscore, workdata)
mod4cdi<- lm(Log.CDI~sex+PVTscore, workdata)
mod5cdi<- lm(Log.CDI~sex+MOTedu, workdata)
anova(mod1cdi)#ct accounts for variance on its own
anova(mod3cdi,mod4cdi)#ct+s+p is better than just s+p
anova(mod2cdi,mod5cdi)#ct+s+m is NOT better than just s+m
```

```{r Testing CDI comp data}

workdata <- workdata %>%
  mutate(outlier.CDIcomp = (Log.CDIcomp > 
                      (mean(Log.CDIcomp, na.rm=T) + 
                         3*(sd(Log.CDIcomp, na.rm=T))) |
                      Log.CDIcomp <
                      (mean(Log.CDIcomp, na.rm=T) - 
                         3*(sd(Log.CDIcomp, na.rm=T)))))

workdata %>% filter(outlier.CDIcomp == T) # 34 is an outlier

workdata.comp <- workdata %>% filter(subj != "34")

shapiro.test(workdata.comp$CDIcomp) # not normal
shapiro.test(workdata.comp$Log.CDIcomp) # not normal


workdata.comp %>% summarize(M = mean(CDIcomp, na.rm=T), 
                                              SD = sd(CDIcomp, na.rm=T))

workdata.comp %>% group_by(caretype) %>% summarize(M = mean(CDIcomp, na.rm=T), 
                                              SD = sd(CDIcomp, na.rm=T))

modelCDIcomp <- lm(Log.CDIcomp ~ sex + PVTscore + MOTedu + caretype, workdata.comp) 
modelCDIcomp.null <- lm(Log.CDIcomp ~ sex + PVTscore + MOTedu, workdata.comp)

anova(modelCDIcomp, modelCDIcomp.null) # n.s.
```


```{r Testing types data}
shapiro.test(workdata$CHItypes) # not normal
shapiro.test(workdata$Log.Types) # normal

workdata <- workdata %>%
  mutate(outlier.types = (Log.Types > 
                      (mean(Log.Types, na.rm=T) + 
                         3*(sd(Log.Types, na.rm=T))) |
                      Log.Types <
                      (mean(Log.Types, na.rm=T) - 
                         3*(sd(Log.Types, na.rm=T)))))

workdata %>% filter(outlier.types == T) # no outliers

workdata %>% summarize(M = mean(CHItypes, na.rm=T), 
                                              SD = sd(CHItypes, na.rm=T))

workdata %>% group_by(caretype) %>% summarize(M = mean(CHItypes, na.rm=T), 
                                              SD = sd(CHItypes, na.rm=T))


mod1types<- lm(Log.Types~caretype, workdata)
mod2types<- lm(Log.Types~caretype+sex+MOTedu, workdata)
mod3types<- lm(Log.Types~caretype+sex+PVTscore, workdata)
mod4types<- lm(Log.Types~sex+PVTscore, workdata)
mod5types<- lm(Log.Types~sex+MOTedu, workdata)
anova(mod1types)#ct accounts for var on its own
anova(mod3types,mod4types)#adding ct to s&pvt helps
anova(mod2types, mod5types)#adding ct to s&motedu helps


```


```{r Testing tokens data}
shapiro.test(workdata$CHItokens) # not normal
shapiro.test(workdata$Log.Tokens) # normal

workdata <- workdata %>%
  mutate(outlier.tokens = (Log.Tokens > 
                      (mean(Log.Tokens, na.rm=T) + 
                         3*(sd(Log.Tokens, na.rm=T))) |
                      Log.Tokens <
                      (mean(Log.Tokens, na.rm=T) - 
                         3*(sd(Log.Tokens, na.rm=T)))))

workdata %>% filter(outlier.tokens == T) # no outliers

workdata %>% summarize(M = mean(CHItokens, na.rm=T), 
                                              SD = sd(CHItokens, na.rm=T))

workdata %>% group_by(caretype) %>% summarize(M = mean(CHItokens, na.rm=T), 
                                              SD = sd(CHItokens, na.rm=T))

mod1tokens<- lm(Log.Tokens~caretype, workdata)
mod2tokens<- lm(Log.Tokens~caretype+sex+MOTedu, workdata)
mod3tokens<- lm(Log.Tokens~caretype+sex+PVTscore, workdata)
mod4tokens<- lm(Log.Tokens~sex+PVTscore, workdata)
mod5tokens<- lm(Log.Tokens~sex+MOTedu, workdata)
anova(mod1tokens)#ct accounts for var on its own
anova(mod3tokens,mod4tokens)#adding ct to s&pvt helps
anova(mod2tokens, mod5tokens)#adding ct to s&motedu helps

```


```{r post-hoc tests}

Home <- subset(workdata, caretype == "Home-only")
Mixed <- subset(workdata, caretype == "Mixed-care")
Full <- subset(workdata, caretype == "Care-only")

# CDI

t.test(Home$Log.CDI, Mixed$Log.CDI) # p<.05
t.test(Full$Log.CDI, Mixed$Log.CDI) # ns
t.test(Home$Log.CDI, Full$Log.CDI) # ns

cohen.d(Home$Log.CDI, Mixed$Log.CDI, na.rm=T, pooled=T) # -.88, large
cohen.d(Full$Log.CDI, Mixed$Log.CDI, na.rm=T, pooled=T) # -.6, medium
cohen.d(Full$Log.CDI, Home$Log.CDI, na.rm=T, pooled=T) # .35, small


# Types

t.test(Home$Log.Types, Mixed$Log.Types) # p=.02
t.test(Full$Log.Types, Mixed$Log.Types) # p=.001
t.test(Home$Log.Types, Full$Log.Types) # ns


cohen.d(Home$Log.Types, Mixed$Log.Types, na.rm=T, pooled=T) # -.98, large
cohen.d(Full$Log.Types, Mixed$Log.Types, na.rm=T, pooled=T) # -1.33, large
cohen.d(Full$Log.Types, Home$Log.Types, na.rm=T, pooled=T) # -.36, small

# Tokens

t.test(Home$Log.Tokens, Mixed$Log.Tokens) # p=.009
t.test(Full$Log.Tokens, Mixed$Log.Tokens) # p<.001
t.test(Home$Log.Tokens, Full$Log.Tokens) # n.s.

cohen.d(Home$Log.Tokens, Mixed$Log.Tokens, na.rm=T, pooled=T) # -1.06, large
cohen.d(Full$Log.Tokens, Mixed$Log.Tokens, na.rm=T, pooled=T) # -1.6, large
cohen.d(Full$Log.Tokens, Home$Log.Tokens, na.rm=T, pooled=T) # -.5, small

```

```{r data for t tests in JASP}
workdata %>% filter(caretype != "Home-only") %>% write.csv("data/noHomebayes.csv")
workdata %>% filter(caretype != "Mixed-care") %>% write.csv("data/noMixedbayes.csv")
workdata %>% filter(caretype != "Care-only") %>% write.csv("data/noFullbayes.csv")
```

```{r, Figure 1}
Figure1a <- ggplot(workdata, mapping=aes(y=Log.CDI, x=caretype, fill=caretype)) + 
  geom_violin(alpha=0.5) + 
  stat_summary(fun.y=mean, geom = "point", aes(group = subj), shape=1, size=3, colour = "black", position = position_jitter(.03)) +
  stat_summary(fun.data=mean_cl_boot, geom = "errorbar", colour='red',  width =.2, size = .8) +
  stat_summary(fun.y=mean, geom = "point", colour='white',  shape=24, fill = "red", size = 5) +
  scale_x_discrete(name=element_blank(),limits=c("Home-only", "Mixed-care", "Care-only"), labels=c("Home-only", "Mixed-care", "Care-only")) +
  scale_y_discrete(limits = c(1.791759, 3.931826, 5.942799),
                     labels = c("5", "50", "380")) +
  guides(fill=guide_legend(title=NULL)) + 
  guides(fill=FALSE, colour = F) +
  ylab('Productive vocabulary at 1;5') + xlab('') +
  theme_bw(base_size=24) +
  theme(axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12),
        legend.title=element_blank())+
  scale_color_manual(values = c("black","grey")) +
  ggtitle("CDI")
#plot(Figure1a)
Figure1b <- ggplot(workdata, mapping=aes(y=Log.Types, x=caretype, fill=caretype)) + 
  geom_violin(alpha=0.5) + 
  stat_summary(fun.y=mean, geom = "point", aes(group = subj), shape=1, size=3, colour = "black", position = position_jitter(.03)) +
  stat_summary(fun.data=mean_cl_boot, geom = "errorbar", colour='red',  width =.2, size = .8) +
  stat_summary(fun.y=mean, geom = "point", colour='white',  shape=24, fill = "red", size = 5) +
  scale_x_discrete(name=element_blank(),limits=c("Home-only", "Mixed-care", "Care-only"), labels=c("Home-only", "Mixed-care", "Care-only")) +
  scale_y_discrete(limits = c(0, 0.6931472, 1.94591, 3.044522, 4.0073332),
                     labels = c("0", "1", "6", "20", "54")) +
  guides(fill=guide_legend(title=NULL)) + 
  guides(fill=FALSE, colour = F) +
  ylab('') + xlab('') +
  theme_bw(base_size=24) +
  theme(axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12),
        legend.title=element_blank())+
  scale_color_manual(values = c("black","grey")) +
  ggtitle("Noun Types")

Figure1c <- ggplot(workdata, mapping=aes(y=Log.Tokens, x=caretype, fill=caretype)) + 
    geom_violin(alpha=0.5) + 
  stat_summary(fun.y=mean, geom = "point", aes(group = subj), shape=1, size=3, colour = "black", position = position_jitter(.03)) +
  stat_summary(fun.data=mean_cl_boot, geom = "errorbar", colour='red',  width =.2, size = .8) +
  stat_summary(fun.y=mean, geom = "point", colour='white',  shape=24, fill = "red", size = 5) +
  scale_x_discrete(name=element_blank(),limits=c("Home-only", "Mixed-care", "Care-only"), labels=c("Home-only", "Mixed-care", "Care-only")) +
    scale_y_discrete(limits = c(0, 1.94591, 4.0073332, 5.620401),
                     labels = c("0", "6", "54", "275")) +
  guides(fill=guide_legend(title=NULL)) + 
  guides(fill=FALSE, colour = F) +
  ylab('') + xlab('') +
  theme_bw(base_size=24) +
  theme(axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12),
        legend.title=element_blank()) +
  scale_color_manual(values = c("black","grey")) +
  ggtitle("Noun Tokens")
#plot(Figure1c)
multiplot(Figure1a, Figure1b, Figure1c, cols=3)
```


```{r, Figure 1 not log-transformed}
Figure1a <- ggplot(workdata, mapping=aes(y=CDI, x=caretype, fill=caretype)) + 
  geom_violin(alpha=0.5) + 
  stat_summary(fun.y=mean, geom = "point", aes(group = subj), shape=1, size=3, colour = "black") +
  stat_summary(fun.data=mean_cl_boot, geom = "errorbar", colour='red',  width =.2, size = .8) +
  stat_summary(fun.y=mean, geom = "point", colour='white',  shape=24, fill = "red", size = 5) +
  scale_x_discrete(name=element_blank(),limits=c("Home-only", "Mixed-care", "Care-only"), labels=c("Home-only", "Mixed-care", "Care-only")) +
  guides(fill=guide_legend(title=NULL)) + 
  guides(fill=FALSE, colour = F) +
  ylab('Productive vocabulary at 1;5') + xlab('') +
  theme_bw(base_size=24) +
  theme(axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12),
        legend.title=element_blank())+
  scale_color_manual(values = c("black","grey")) +
  ggtitle("CDI")
Figure1b <- ggplot(workdata, mapping=aes(y=CHItypes, x=caretype, fill=caretype)) + 
  geom_violin(alpha=0.5) + 
  stat_summary(fun.y=mean, geom = "point", aes(group = subj), shape=1, size=3, colour = "black") +
  stat_summary(fun.data=mean_cl_boot, geom = "errorbar", colour='red',  width =.2, size = .8) +
  stat_summary(fun.y=mean, geom = "point", colour='white',  shape=24, fill = "red", size = 5) +
  scale_x_discrete(name=element_blank(),limits=c("Home-only", "Mixed-care", "Care-only"), labels=c("Home-only", "Mixed-care", "Care-only")) +
  guides(fill=guide_legend(title=NULL)) + 
  guides(fill=FALSE, colour = F) +
  ylab('') + xlab('') +
  theme_bw(base_size=24) +
  theme(axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12),
        legend.title=element_blank())+
  scale_color_manual(values = c("black","grey")) +
  ggtitle("Noun Types")

Figure1c <- ggplot(workdata, mapping=aes(y=CHItokens, x=caretype, fill=caretype)) + 
    geom_violin(alpha=0.5) + 
  stat_summary(fun.y=mean, geom = "point", aes(group = subj), shape=1, size=3, colour = "black") +
  stat_summary(fun.data=mean_cl_boot, geom = "errorbar", colour='red',  width =.2, size = .8) +
  stat_summary(fun.y=mean, geom = "point", colour='white',  shape=24, fill = "red", size = 5) +
  scale_x_discrete(name=element_blank(),limits=c("Home-only", "Mixed-care", "Care-only"), labels=c("Home-only", "Mixed-care", "Care-only")) +
  guides(fill=guide_legend(title=NULL)) + 
  guides(fill=FALSE, colour = F) +
  ylab('') + xlab('') +
  theme_bw(base_size=24) +
  theme(axis.title.y = element_text(size = 15),
        axis.text.x = element_text(size = 12),
        legend.title=element_blank()) +
  scale_color_manual(values = c("black","grey")) +
  ggtitle("Noun Tokens")

multiplot(Figure1a, Figure1b, Figure1c, cols=3)
```

# Supplementary data

```{r Work status group size for supplementary materials}
workdata %>% group_by(MOTwork6, sex) %>% tally()

workdata %>% group_by(MOTwork18, sex) %>% tally()

workdata %>% group_by(MOTwork6, MOTedu) %>% tally()

workdata %>% group_by(MOTwork18, MOTedu) %>% tally()

workdata %>% group_by(MOTwork18, family_size) %>% tally()

workdata %>% group_by(MOTwork6, family_size) %>% tally()

workdata %>% group_by(MOTwork18, siblings) %>% tally()

workdata %>% group_by(MOTwork6, siblings) %>% tally()

workdata %>% group_by(FATwork6, MOTwork6) %>% tally()

workdata %>% group_by(FATwork18, MOTwork18) %>% tally()

FigS1a <- ggplot(workdata, aes(x=MOTworkhours6, fill=MOTwork6)) +
    geom_histogram(binwidth = 1.5) +
  theme_bw() +
  theme(legend.title=element_blank(),
        legend.position=c(.80, .85)) +
  xlab("Hours worked at 6 months") +
    coord_cartesian(ylim=c(0, 20))


FigS1b <- ggplot(workdata, aes(x=MOTworkhours18, fill=MOTwork18)) +
    geom_histogram(binwidth = 1.5) +
  theme_bw() +
  theme(legend.position = "none") +
  xlab("Hours worked at 18 months") +
  coord_cartesian(ylim=c(0, 20))

multiplot(FigS1a, FigS1b, cols=2)
```

```{r Distribution of working hours across groups 6m}
FigS2a <- ggplot(workdata) +
  geom_point(mapping=aes(x=MOTworkhours6, y=CHItypes, colour=MOTwork6, shape=MOTwork6), position=position_jitter(0.1), size=3) +
  scale_colour_manual(name = "Care Schedule", labels = c("Home-only", "Mixed-care", "Care-only"), values = c("#F8766D", "#00BA38", "#619CFF")) +
  scale_shape_manual(name = "Care schedule", labels = c("Home-only", "Mixed-care", "Care-only"), values=c(0,1,2)) +
  geom_smooth(mapping=aes(x=MOTworkhours6, y=CHItypes), method='lm') +
  guides(fill=guide_legend(title=NULL)) +
  ylab('Word types produced at 1;5') + xlab("Mothers' weekly work hours at 6m") +
  theme_bw(base_size=12) +
  theme(legend.position = 'none')

FigS2b <- ggplot(workdata) +
  geom_point(mapping=aes(x=MOTworkhours6, y=CHItokens, colour=MOTwork6, shape=MOTwork6), position=position_jitter(0.1), size=3) +
  scale_colour_manual(name = "Care Schedule", labels = c("Home-only", "Mixed-care", "Care-only"), values = c("#F8766D", "#00BA38", "#619CFF")) +
  scale_shape_manual(name = "Care schedule", labels = c("Home-only", "Mixed-care", "Care-only"), values=c(0,1,2)) +
  geom_smooth(mapping=aes(x=MOTworkhours6, y=CHItokens), method='lm') +
  guides(fill=guide_legend(title=NULL)) +
  ylab('Word tokens produced at 1;5') + xlab("Mothers' weekly work hours at 6m") +
  theme_bw(base_size=12) +
  theme(legend.position = 'none')

multiplot(FigS2a, FigS2b, cols=2)

```

```{r Distribution of working hours across groups 18m}
FigS3a <- ggplot(workdata) +
  geom_point(mapping=aes(x=MOTworkhours18, y=CHItypes, colour=MOTwork18, shape=MOTwork18), position=position_jitter(0.1), size=3) +
  scale_colour_manual(name = "Care Schedule", labels = c("Home-only", "Mixed-care", "Care-only"), values = c("#F8766D", "#00BA38", "#619CFF")) +
  scale_shape_manual(name = "Care schedule", labels = c("Home-only", "Mixed-care", "Care-only"), values=c(0,1,2)) +
  geom_smooth(mapping=aes(x=MOTworkhours18, y=CHItypes), method='lm') +
  guides(fill=guide_legend(title=NULL)) +
  ylab('Word types produced at 1;5') + xlab("Mothers' weekly work hours at 18m") +
  theme_bw(base_size=12) +
  theme(legend.position = 'none')

FigS3b <- ggplot(workdata) +
  geom_point(mapping=aes(x=MOTworkhours18, y=CHItokens, colour=MOTwork18, shape=MOTwork18), position=position_jitter(0.1), size=3) +
  scale_colour_manual(name = "Care Schedule", labels = c("Home-only", "Mixed-care", "Care-only"), values = c("#F8766D", "#00BA38", "#619CFF")) +
  scale_shape_manual(name = "Care schedule", labels = c("Home-only", "Mixed-care", "Care-only"), values=c(0,1,2)) +
  geom_smooth(mapping=aes(x=MOTworkhours18, y=CHItokens), method='lm') +
  guides(fill=guide_legend(title=NULL)) +
  ylab('Word tokens produced at 1;5') + xlab("Mothers' weekly work hours at 18m") +
  theme_bw(base_size=12) +
  theme(legend.position = 'none')

multiplot(FigS3a, FigS3b, cols=2)

```

```{r Correlation between mothers work hours and productive vocabulary and household Income}

cor.test(workdata$HouseholdIncome18, workdata$HouseholdIncome6, method = 'pearson') # Significant
cor.test(workdata$MOTworkhours6, workdata$HouseholdIncome6, method = 'pearson') # Significant
cor.test(workdata$MOTworkhours18, workdata$HouseholdIncome6, method = 'pearson') # Significant
cor.test(workdata$Log.Types, workdata$HouseholdIncome6, method = 'pearson') # Not Significant
cor.test(workdata$Log.Tokens, workdata$HouseholdIncome6, method = 'pearson') # Not Significant

cor.test(workdata$MOTworkhours6, workdata$HouseholdIncome18, method = 'pearson') # Significant
cor.test(workdata$MOTworkhours18, workdata$HouseholdIncome18, method = 'pearson') # Marginal
cor.test(workdata$Log.Types, workdata$HouseholdIncome18, method = 'pearson') # Not Significant
cor.test(workdata$Log.Tokens, workdata$HouseholdIncome18, method = 'pearson') # Not Significant


cor.test(workdata$MOTworkhours18, workdata$MOTworkhours6, method = 'pearson') # Significant
cor.test(workdata$Log.Types, workdata$MOTworkhours6, method = 'pearson') # Marginal
cor.test(workdata$Log.Tokens, workdata$MOTworkhours6, method = 'pearson') # Not Significant

cor.test(workdata$Log.Types, workdata$MOTworkhours18, method = 'pearson') # Marginal
cor.test(workdata$Log.Tokens, workdata$MOTworkhours18, method = 'pearson') # Not Significant

cor.test(workdata$Log.Tokens, workdata$Log.Types, method = 'pearson') # Not Significant

```
